\begin{acknowledgments}

The journey towards completing my Ph.D. has been an incredible one, filled with both triumphs and challenges. I am grateful to the many individuals who have supported me along the way, including professors, colleagues, friends, family, and participants. Their unwavering support and encouragement have played a crucial role in shaping me as a better researcher and as an individual.

First and foremost, I would like to express my deepest gratitude to my Ph.D. advisor, Prof.~Shengdong Zhao (Shen), for his guidance throughout this journey. He has not only imparted valuable knowledge and research skills but also instilled in me the importance of critical thinking and pushing boundaries. I am also grateful for his support and guidance as a colleague. 

I would like to extend my thanks to all the members of the NUS-HCI lab, including but not limited to Devangini Patel, Debjyoti Ghosh, Zhang Shan, Hyeongcheol Kim, Xinke Wu, Smitha Sheshadri, Ashwin Ram, Felicia Tan, Shardul Sapkota, Cai Runze, Samantha Chan, Chloe Haigh, Yang (Sonne) Chen, Zhang Yue, Zhou Chen, Katherine Fennedy, and Yu Jiahan. I am also thankful to my collaborators from other labs, especially Prof. Petr Slovak, for their guidance and contributions to my projects.

During challenging times, particularly at the beginning of my Ph.D. journey and when things did not go as planned, I was fortunate to have friends who provided invaluable support in research and life. I would like to thank Thilina Madusanka Thanthriwatta, Ridwan Shariffdeen, Sunimal Rathnayake, Dhananjaya Wijerathne, Manupa Karunaratne, Ruba Panchendrarajan, Thushara Sandakalum, Dileepa Fernando, Sanka Rasnayaka, Tharindu Miyanawala, Sandareka Wickramanayake, Viraj Herath, Dumith Jayathilaka, Ashan Asmone, Tharindu Madanayake, Supun Tharanga, Dulanga Shashika, Supuli Jayaweera, Gayana Herath, Himeshi De Silva, Nirandika Wanigasekara, Suman Sourav, and Suman Panda for making my time in Singapore truly memorable.


I would also like to express my gratitude to the teaching module coordinators, Prof.~Brian Y. Lim, Prof.~Damith C. Rajapakse, and Prof.~Bimlesh Wadhwa, for their support and patience in helping me learn new topics during my teaching assistant duties. I am also thankful to Prof.~Tan Chuan Hoo, Prof.~Isam Faik, Prof.~Leow Wee Kheng, and Prof.~Abhik Roychoudhury for their guidance in conducting research.

I extend my sincere appreciation to my dissertation committee members, Prof.~Brian Y. Lim,  Prof.~Suranga Nanayakkara, and Prof.~Roger Zimmermann, for their invaluable feedback, which has significantly contributed to the improvement of this thesis.


My heartfelt thanks go to my family, who have been my unwavering support throughout this journey. Despite the challenges I faced, both in life and in research, you were always there for me, regardless of the distance or time. To my \i{amma}, \i{thatha}, \i{malli}, and Rasadi, thank you from the bottom of my heart.


Finally, I would like to express my gratitude to all the individuals whose names may not have been mentioned but have played a vital role in making this Ph.D. journey possible. The contributions of the administration services (e.g., Ms.~Loo Line Fong), technical services, BuildFac services, transport services, accommodation providers, participants, reviewers, cleaning services, and many others have all played a significant role in the success of this journey. Last but certainly not least, I would like to express my deepest gratitude to all the funding agencies that have supported our research. 

Thank you all!


\end{acknowledgments}
